---
title: "twitter"
author: "Brenden Latham"
date: "6/23/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("tidyverse")
library("ggplot2")
library("tidytext")
library("gtools")
```

# Vaccine Sentiment Analysis

## Initial Data
1,998,849 tweets containing the word "vaccine" posted since January 1st, 2020 were scraped from twitter using Twint.

To install and run twint in linux:

conda install twint

twint -s vaccine --since 2020-1-1 --until 2021-6-14 --csv -o vaccine_mentions.csv 

## Would scraping for hashtags yield better results?

The data will be brought into Rstudio and a list of hashtags used in conjunction with the word "vaccine" will be produced in order to determine if hashtags are a better metric for scrapping.

```{r load}
vac_mentions <- read.csv("~/twitter/vaccine_mentions.csv", sep = "\t", encoding = "UTF-8") ## edit path to file
```

```{r}
head(vac_mentions,1)
```

Twint formats all hashtags in a tweet as an array so some cleaning is necessary.

In the first step of cleaning a list in the form of a large character will be made.

```{r tags1}
hashtags <- vac_mentions %>% filter(language == "en") %>% select("hashtags") ## creating data frame with only english hashtags
hashtags <- hashtags[!(hashtags$hashtags=="[]"),] ## deleting rows containing what Twint gives as a NULL value
hashtags <- as.vector(unlist(strsplit(hashtags,",")),mode="list") ## splitting each hashtag into it's own element in a list
hashtags <- str_replace_all(hashtags, "[^[:alnum:]]", "") ## deleting all special characters
hashlist <- unique(hashtags) ## making list of unique hashtags
hashlist <- paste0("#", hashlist) ## putting it in hashtag format
head(hashlist)
```

In the second step a frequency table of hashtag use will be made and saved as a dataframe.

```{r tags2}
hashframe <- hashtags ## createing another list with only hashtags
hashframe <- paste0("#", hashframe) ## putting it in hashtag format
hashframe <- as.data.frame(hashframe) ## converting it back to data frame
hash_count <- plyr::count(hashframe$hashframe) ## creating frequency table
hash_count <- hash_count[order(-hash_count$freq),] ## ordering it by most occurring
head(hash_count)
#write.csv
```

A graph reprsenting this data is shown below.

```{r tagplot}
boxplot(hash_count$freq, main = "frequency of hashtags", xlab = "hashtags", ylab = "frequency")
```

It is evident that there are outliers in the hashtags, with the furthest being a hashtag not necessarily always pertaining to a vaccine.

Now the number of tweets gathered with no hashtags will be shows.

```{r nulltags1}
null_count <- plyr::count(vac_mentions$hashtags) # making frame with NULLS included
head(null_count[order(-null_count$freq),]) # ordering frame to see nuber of NULLS
```

These results reveal that hashtags are not a reliable metric to use when scraping for tweets pertaining to a vaccine, and further scraping using hashtags as keywords would likely result in problems when analysing tweet sentiment.

Using the database of tweets containing "vaccine" has been ruled to be the best method of analyzing the current public opinion on the covid vaccine and vaccines in general.

Analysis will move forward using the dataframe of vaccine mentions.

## User

The use of the word vaccine in relation to user will now be analyzed.

```{r usercount}
user_freq <- plyr::count(vac_mentions$username) # getting frequency of users
head(user_freq[order(-user_freq$freq),]) # ordering by most frequent
```

```{r userplot}
boxplot(user_freq$freq, main = "tweets per user pertaining to vaccine", xlab = "user", ylab = "frequency")
```

The mean will be calculated below.

```{r meantweets}
mean(user_freq$freq)
```

Of twitter users that have tweeted the word "vaccine" one time, the mean amount of tweets containing the word vaccine is approximately 2.28 with the highest outlier being 14,724.

The usernames of the top tweeters indicate bot accounts.

The users will now be split into a large list of individual databases containing all tweets for that user.

```{r}
users_split <- split(vac_mentions,f = vac_mentions$username) # splitting users into individual dataframes within a list
```

The first six tweets from the top tweeter will be shown below

```{r}
head(users_split[["vaccineedgar"]][["tweet"]]) # examining the top tweeter
```

Now some tweets will be shown from a slightly different type of account

```{r}
head(users_split[["001_venom"]][["tweet"]]) # examining another user
```

The links contained in these posts lead to various anti-vaccine videos and images.

There is a noticeable difference between the amount of hashtag spam between these two posts. Level of hashtag spam might be useful in determining the purpose of the tweet

The average number of hashtags used by each user will be evaluated.

```{r}

```

